\Title

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass: intro

This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms. Our SNN model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include heterogeneous delays in addition to weights. In particular, one afferent may be connected with multiple delays. Crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous MLR model by encoding spiking motifs as spatio-temporal patterns. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these heterogeneous delays. %

@@

In this work, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}). The paper is organized as follows. We develop a theoretically defined SNN capable of learning heterogeneous delays from spiking motifs that we first evaluate on a synthetic event-based dataset. We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays. This will allow us to formalize the spiking neuron used to learn the model's parameters and test its effectiveness. We aim to apply this model to a realistic motion detection task which we will first define. Then, we will show how the model can be adapted to solve the motion detection task. In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations. We will also test the network with classical stimuli used in neuroscience, and in particular with frequency or contrast responses.  We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off.  Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations. This will open perspectives for future SNNs.  In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is perfectly fit to be transfered to this type of hardware.

>>>>

This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms, notably by using Heterogeneous Delays in Spiking Neural Networks (HD-SNN). Our HD-SNN model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include heterogeneous delays in addition to weights. In particular, one afferent may be connected with multiple delays. Crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous MLR model~\citep{grimaldi_robust_2022} by explictly encoding spiking motifs as spatio-temporal patterns. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these heterogeneous delays. %

@@

In this work, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}). The paper is organized as follows. We develop a theoretically defined HD-SNN capable of learning heterogeneous delays that we first evaluate on a synthetic event-based dataset. We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays. This will allow us to formalize the spiking neuron used to learn the model's parameters and test its effectiveness. We aim to apply this model to a realistic motion detection task which we will first define. Then, we will show how the model can be adapted to solve the motion detection task. In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations. We will also test the network with classical stimuli used in neuroscience, and in particular with frequency or contrast responses.  We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off.  Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations. This will open perspectives for future SNNs.  In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is perfectly fit to be transfered to this type of hardware and to obtain significant gains in the energy which is used.
